An IFS study of Low-powered Radio Galaxies

### 2019-08-26 paper.tex
Changes from Martin - his notes are in {\bf ...}
 On branch master
 Changes to be committed:
	modified:   paper.tex
	modified:   refs.bib

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[An IFS study of Low-powered Radio Galaxies]{An IFS study of Low-powered Radio Galaxies}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[J. Warren et al.]{
Joshua Warren,$^{1}$\thanks{Contact e-mail: \href{mailto:warrenj@gmx.com}{warrenj@gmx.com}}
%Joshua Warren,$^{1}$\thanks{Contact e-mail: \href{mailto:joshua.warren@physics.ox.ac.uk}{joshua.warren@physics.ox.ac.uk}}
Martin Bureau,$^{1}$
Bernd Hasemann,$^{2}$
Isabella Prandoni,$^{3}$ \newauthor
Francesco Santoro,$^{3}$
Robert Laing,$^{3}$
Paola Parma,$^{3}$
Hans de Ruiter$^{3}$
and \newauthor
Arturo Mignano$^{3}$
\\
$^{1}$Sub-department of Astrophysics, Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK\\
$^{2}$European Southern Observatory, Karl-Schwarzschild-Str. 2, 85748 Garching b. München, Germany\\
$^{3}$INAF - Istituto di Radioastronomia, Via P. Gobetti 101, 40129 Bologna, Italy}

@@

% Abstract of the paper
\begin{abstract}
The most powerful radio galaxies are well studied with a strong consensus in the literature regarding their formations processes. However, they are intrinsically rare objects and not representative of the general population of radio galaxies. Low-powered radio galaxies are comparatively poorly studied with many conflicting viewpoints regarding their origin and the extent of the effect of feedback from them.

We present a complete sample of 11 nearby low-powered radio galaxies, observed with a suite of instruments. This paper focuses on the spatially-resolved stellar kinematics and populations, and ionized gas distribution, kinematics and ionization derived from optical IFS observations%, while other observation such as \ce{^{12}CO(2-1)} from Atacama Large Millimetre/sub-millimetre Array (ALMA) and radio continuum from the Very Large Array (VLA) will be presented in future papers
. We compare the characteristics and properties of our sample to those of ordinary (i.e.\ optically-selected) ETGs using the Atlas$^\text{3D}$ and MASSIVE IFS surveys. These include the fast/slow rotator classification fraction, kinematic substructures, absorption line strength radial gradients, general stellar population properties, ionized gas distribution and kinematics and the dominant source of ionizing photons.

We tentatively find that our radio-selected ETGs contain a higher mass in of ionized gas than expected. However, after accounting for the differences in the mass distribution, we find no further differences between the two samples in the characteristics that can be studied with optical IFU data. We suggest that this is consistent with the notion that low-powered radio galaxies are ordinary ETGs experiencing a brief active phase which maintains their low star-formation rates (also known as maintenance or jet-mode AGN). 
\end{abstract}

@@

\section{Introduction}
	\label{sec:intro}
	It is becoming increasingly clear that active galactic nuclei (AGN) can be separated into two different types, radiative and jet mode AGN \citep[e.g.][]{Antonucci2012, Heckman2014}, raising speculation about differing mechanisms and possibly fuel sources for each. The former is thought to be the powerful engine that initially and rapidly shuts off star formation in massive early-type galaxies (ETGs; e.g. \citealt{Thomas2005, Thomas2010}), either by removing the cold gas from the galaxy entirely, heating it such that it joins the hot X-ray gas often present in massive galaxies \citep[e.g.][]{OSullivan2001}, or rendering it unusable for star formation (such as by stabilising it against collapse; e.g.\ \citealt{Martig2009}). After a single radiative-mode episode, the AGN is thought to subside to become jet-dominated, repeatedly reheating the surrounding interstellar medium (ISM) to counter any cooling that might allow the cold gas reservoir to replenish (and subsequently form stars).

	There is a strong consensus within the scientific literature that the most powerful radio galaxies are the result of gas-rich, major mergers causing large amounts of material to fall on to the black hole at the centre of the resulting galaxy, thus producing an immensely powerful AGN which includes a powerful radio jet \citep[e.g.][]{Malin1983, Quillen1992, Lim2000}. However, such galaxies are extremely rare and are not representative of the general radio galaxy population. Most radio galaxies are low-powered, and their basic properties and origins are poorly constrained. One of the most popular suggestions is that they are bona-fide detections of jet-mode AGN \citep[e.g.][]{Heckman2014}.%, a class of AGN thought to maintain the dearth of star formation in early-type galaxies (ETGs), after most of the fuel for star formation has been removed by an earlier event. 
	These radio galaxies are known to be dusty with dust lanes having a tendency to be perpendicular to the jets \citep[e.g.][]{DeRuiter2002, VerdoesKleijn2005}. They have little ionized gas in the ISM \citep[e.g.][]{Sarzi2005}, but often have significant quantities of hot X-ray gas (e.g. \citealt{Canizares1987}; the cold phase of the ISM of radio galaxies will be examined in Ruffa et.al.\ in prep.). If the low star-formation rates typically seen in massive ETGs are preserved by jet-mode AGN, then all such ETGs must regularly go through similar active phases, and therefore low-powered radio galaxies should be much the same as their radio-quiet (quiescent phase) ETG counterparts.

	To test this hypothesis, a complete sample of 11 nearby low-powered radio galaxies was constructed by \citet{Prandoni2010} and observed with a suite of instruments. Here, we present the spatially-resolved stellar kinematics and populations, and ionized gas distribution, kinematics and ionization derived from integral-field spectroscopic (IFS) observations with the Visible Multi-object Spectrograph (VIMOS) and archival data from the Multi-unit Spectroscopic Explorer (MUSE), both on the Very Large Telescope (VLT). We compare the characteristics and properties (accessible with $V$-band IFS data) of our sample of low-power radio galaxies to those of ordinary (i.e.\ optically-selected) ETGs from the Atlas$^\text{3D}$ and MASSIVE IFS surveys. We hereafter refer to the combined Atlas$^\text{3D}$ \citep{Cappellari2011} and MASSIVE \citep{Ma2014} sample as the A+M sample.

	In Section \ref{sec:samp} of this paper we set out the compete sample with our observations and data-reduction methods detailed in Section \ref{sec:obs}. Our data-analysis methods and routines are detailed in Section \ref{sec:analysis} (the maps of our sample produced with these methods are shown in Appendix \ref{sec:maps}). We present our findings regarding the stellar kinematics, absorption line strengths, stellar populations, ionized gas (distribution, kinematics and ionization) in Sections \ref{sec:StarKine}-\ref{sec:gas}, respectively. Finally, we discuss our results and conclude in Section \ref{sec:discuss}.

\section{Sample Description}
	\label{sec:samp}
	The Parkes 2.7 GHz survey \citep{Ekers1989} was used as a parent sample, as it contains radio sources (down to a flux density limit of 250\,mJy at 2.7\,GHz) with an optical counterpart (with an $V$-band apparent magnitude of $m_V \le 17.0$). Host galaxies with a redshift greater than 0.03 were discarded. It is worth noting that both the radio and optical limits are apparent measurements, and as such this is not a volume-limited sample.

	\begin{table*}
		\centering
	% \begin{threeparttable}
		\caption{Key characteristics of the Southern Sample galaxies. Col.\,1: host galaxy name. Col.\,2: radio source name from Parkes Southern Radio Source Catalogue. We hereafter refer to the last galaxy by its PKS name only. Col.\,3: redshift. Col.\,4: National Radio Astronomy Observatory (NRAO) VLA Sky Survey (NVSS) 1.4 GHz flux density \citep{Condon1998}. Col.\,5: Two Micron All Sky Survey (2MASS) $K$-band apparent magnitude (with errors of $\pm 0.025$ mag; \citealt{Skrutskie2006}). Col.\,6: dust morphologies from (a) \citet{Govoni2000}, (b) \citet{Lauer2005}, (c) \citet{Bettoni2001}, (d) \citet{Sandage1979}, (e) \citet{Colbert2001}.}
		\label{tab:sample}
		\begin{tabular}{l c c c c l}
			\hline
			\hline
			Host galaxy	& Radio source 	& Redshift	& $S_\text{1.4\,GHz}$	& $m_K$ & Dust morphology\\
						& (PKS) 		& 			& (mJy) 			& (mag)	&\\
			\hline 
			ESO 443-G024 & 1258-321 	& 0.01703	& $1387 \pm 45$		& 8.51 & no dust$^{a}$	\\ 
			IC 1459 	& 2254-367 		& 0.00565 	& $1280 \pm 45$		& 6.81 & dust lane$^{b}$	\\
			IC 1531 	& 0007-325 		& 0.02553 	& \leavevmode\phantom{0}$388 \pm 12$		& 9.55 & --					\\
			IC 4296		& 1333-\leavevmode\phantom{0}33 		& 0.01248 	& $1342 \pm 41$		& 7.50 & edge-on disc$^{b}$ \\
			NGC 612 	& 0131-\leavevmode\phantom{0}36 		& 0.02954 	& \leavevmode\phantom{0}$585 \pm 18$		& 9.58 & dust lane$^{c}$	\\
			NGC 1316 	& 0320-\leavevmode\phantom{0}37 & 0.00591 	& \leavevmode\phantom{0}$255 \pm 10$		& 5.59 & dust patches$^{b}$ \\
			NGC 1399 	& 0336-\leavevmode\phantom{0}35 & 0.00472 	& \leavevmode\phantom{0}$209 \pm \leavevmode\phantom{0}7$	& 6.31 & no dust$^{b}$	\\
			NGC 3100 	& 0958-314 		& 0.00879 	& \leavevmode\phantom{0}$530 \pm 16$		& 8.08 & dust lane$^{d}$	\\
			NGC 3557 	& 1107-372 		& 0.01016 	& \leavevmode\phantom{0}$484 \pm 16$		& 7.20 & face-on disc$^{b}$\\
			NGC 7075 	& 2128-388 		& 0.01819 	& \leavevmode\phantom{0}$837 \pm 28$		& 9.56 & --					\\
			--			& 0718-\leavevmode\phantom{0}34 		& 0.02897 	& $1119 \pm 41$		& 9.97 & dust patches$^{e}$ \\
			\hline
			\hline
		\end{tabular}
	% 	\begin{tablenotes}
	% 	\footnotesize
	% 	\note Col.\,1: host galaxy name. Col.\,2: radio source name from Parkes Southern Radio Source Catalogue. Col.\,3: redshift. Col.\,4: National Radio Astronomy Observatory (NRAO) Very Large Array (VLA) Sky Survey (NVSS) 1.4 GHz flux density \citep{Condon1998}. Col.\,5: Two Micron All Sky Survey (2MASS) $K$-band apparent magnitude (with errors of $\pm 0.025$ mag; \citealt{Skrutskie2006}). Col.\,6: dust morphologies from (a) \citet{Govoni2000}, (b) \citet{Lauer2005}, (c) \citet{Bettoni2001}, (d) \citet{Sandage1979}, (e) \citet{Colbert2001}. 
	% 	\item We hereafter refer to the last galaxy by its PKS name only.
	% 	\end{tablenotes}
	% \end{threeparttable}
	\end{table*}

	These criteria lead to a sample of 11 galaxies hereafter referred to as the Southern Sample. The key observed characteristics are summarised in Table \ref{tab:sample}. All galaxies have a FR I radio morphology according to the \citet{Fanaroff1974} scheme.

	The sample was initially observed with the Atacama Pathfinder Experiment \citep[APEX; ][]{Gusten2006}, with detections claimed for the \ce{^{12}CO(2-1)} transition for all galaxies. Some galaxies had extremely large line widths (up to a full-width half-maximum of $904 \, \mathrm{km \, s^{-1}}$ for IC 4296), but most showed a flat or double-horn spectrum indicative of ordered rotation \citep{Prandoni2012}.

\section{Observing strategy and Data Reduction}
	\label{sec:obs}
	The sample has been observed with a suit of instruments spanning the radio to optical wavelengths (archival images from the \textit{Chandra space telescope} are also available for some of the sample). This paper focuses on the optical IFS observations made with VIMOS, mounted on the VLT in Paranal \citep{LeFevre2003}. All observations were taken with a spatial sampling of $0\farcs67$, using the high-resolution blue grism yielding a wavelength range of 3700-5520\,\AA\ with a spectral resolution of 1440 and sampling of 0.71\,\AA\,$\mathrm{pix^{-1}}$. Each object was observed with a total integration time of $\approx 100$\,min on-source, equally spread over three observing blocks (OBs). Each OB also contained all the necessary observation for standard IFS data reduction.

	VIMOS has several well-known though poorly-understood technical issues. These include several low transmission (bad) fibres, strong flexure, contamination from adjacent spectra on the CCD (cross-talk) and large differences in sensitivity across its 4 quadrants. To address these issues, we adopted a data-reduction pipeline produced using \textsc{Py3D}, a suite of programs based on the \textsc{python} versions of those developed for the Calar Alto Legacy Integral Field spectroscopy Area survey \citep[CALIFA;][]{Sanchez2012, Husemann2013} but later updated for VIMOS by \citet{Husemann2014}. \textsc{Py3D} applies all standard IFS data-reductions steps: bias subtraction, spectrum identification, flatfielding and wavelength calibration. In addition, the data was flux calibrated using publicly available observations of the spectro-photometric standard star Feige 110 provided by the European Southern Observatory (ESO). All steps include methods designed to account for the bad fibres, flexure and cross-talk within VIMOS.

	Following this, we noted that the cubes were still not properly calibrated. Three main issues remain: the quadrants have different intensities, clearly showing uncorrected throughput differences; diagonal intensity stripes are present in the reconstructed images at all wavelengths; and spectral features are observed \citep{Jullo2008} that are visually similar to fringe patterns caused by interference within the CCD between incident light and light reflected from the interfaces between layers of the CCD materials. \citet{Lagerholm2012} present ad-hoc methods to correct for these issues, which we here adopt. These additional steps mean that the datacubes are not perfectly flux-calibrated, but all the corrections are multiplicative and thus will not effect equivalent width or line ratio measurements. From comparisons to the MUSE data (see Section \ref{subsec:MUSE}), we then estimate the flux calibration of the resulting datacubes.

	These method are not completely effective and some artefacts are still present in the data, however in most cases the data is of sufficient quality for our purposes. For the worst effected datacubes, we have archival MUSE observations (see Section \ref{subsec:MUSE}).

	The variance spectra are propagated throughout the data-reduction pipeline (including these ad-hoc corrections) and are square-rooted at this point, to be used as noise inputs in the following analyses (see Section \ref{sec:analysis}).

	\subsection{MUSE Archival Data}
		\label{subsec:MUSE}
		Four of the sample galaxies are in the MUSE archive: IC 1459, IC 4296, NGC 1316 and NGC 1399. We include these in our analysis as they have better (spatial and spectral) resolution and sampling, a larger field-of-view and larger wavelength range than the VIMOS data.

		We used the pre-reduced (Phase 3) datacubes from ESO. These were generally of sufficient quality for our purpose, except in both IC 1459 and IC 4296 the sky appears to have been over-subtracted. To remove this over-subtraction, we developed our own pseudo-sky subtraction routine. A median spectrum was taken from four $20 \times 20$\,spaxel regions, one in each spatial corner of the ESO reduced cube. After checking that no stellar continuum could be fit to the median sky spectrum (i.e.\ that very little light from the galaxy is contaminating the pseudo-sky regions), this median sky spectrum was subtracted from each spaxel in the cube.

		Finally, we trimmed all the MUSE cubes to the central $30\arcsec \times 30\arcsec$ ($150 \times 150$\,spaxels) only, to (a) avoid the regions used for the pseudo-sky spectrum and (b) reduce the computing resources required for spatial binning (see Section \ref{sec:analysis}). In any case, the bins would likely be so large in the outer parts as to be effectively useless.

\section{Data Analysis}
	\label{sec:analysis}
	Both VIMOS and MUSE datasets are spatially binned using Voronoi binning\footnote{http://www-astro.physics.ox.ac.uk/~mxc/software/} \citep{Cappellari2003} and from this point on are treated almost identically, the only difference (other than the different spectral ranges and resolutions) being that the MUSE datacubes are spatially binned to a higher signal-to-noise ratio (S/N). For our purposes, we define the `signal' and `noise' of each spaxel as the median value of its spectrum and noise spectrum, respectively. We require a S/N of 30 for all VIMOS datacubes and 50 for the IC 1459 and IC 4296 MUSE datacubes. The NGC 1316 and NGC 1399 MUSE datacubes were binned to a S/N of 50 for the analysis of the stellar kinematics and 100 for the analysis of the emission line kinematics and stellar populations. These target S/N were chosen in order to be as low as possible, while still returning meaningful information (as assessed by eye).

	Next we implement a routine to find optimal initial values of the redshift and velocity dispersion for each galaxy, to be used when fitting individual bins, by means of iterative fits to the spatially collapsed (i.e. summed across both spatial dimensions of the datacube) spectrum. We use the penalized-fitting (\textsc{pPXF}) routine\footnotemark[1] of \citet{Cappellari2004} and \citet{Cappellari2016a} with the Medium-resolution Isaac Newton Telescope (INT) Library of Empirical Spectra (MILES) library \citep{Sanchez-Blazquez2006, Falcon-Barroso2011a} to find the best-fitting line-of-sight stellar velocity distribution (LOSVD; as parametrised by the Gaussian parameters $v$, the mean velocity and $\sigma$, the velocity dispersion). The final mean velocity of a given galaxy, is used as its precise spectroscopic redshift and is given in Table \ref{tab:sample}. Beyond this point, each bin is analysed independently.

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[IFS Study of Low-power Radio Galaxies]{An Integral-field
  Spectroscopic Study of Low-power Radio Galaxies}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[J. Warren et al.]{
  Joshua Warren,$^{1}$\thanks{Contact e-mail:
    \href{mailto:warrenj@gmx.com}{warrenj@gmx.com}}
%Joshua Warren,$^{1}$\thanks{Contact e-mail: \href{mailto:joshua.warren@physics.ox.ac.uk}{joshua.warren@physics.ox.ac.uk}}
  Martin Bureau,$^{1,2}$ Bernd Hasemann,$^{3}$ Isabella
  Prandoni,$^{4}$ \newauthor Francesco Santoro,$^{4}$ Robert
  Laing,$^{4}$ Paola Parma,$^{4}$ Hans de Ruiter$^{4}$ and \newauthor
  Arturo Mignano$^{4}$
  \\
  $^{1}$Sub-department of Astrophysics, Department of Physics,
  University of Oxford, Denys Wilkinson Building, Keble Road, Oxford
  OX1 3RH, UK\\
  $^{2}$Yonsei Frontier Lab and Department of Astronomy, Yonsei
  University, 50 Yonsei-ro, Seodaemun-gu, Seoul 03722, Republic of
  Korea\\
  $^{3}$European Southern Observatory, Karl-Schwarzschild-Str.\ 2,
  85748 Garching b.\ M{\"u}nchen, Germany\\
  $^{4}$INAF - Istituto di Radioastronomia, Via P.\ Gobetti 101, 40129
  Bologna, Italy\\
  $^*${\bf Should you/we add Ilaria Ruffa and/or Rosita Paladino
    and/or Timothy Davis as authors (they are all on the first/ALMA
    paper)?}\\{\bf I feel it is a little strange that Ilaria is not (so I
    would argue we should add her), I am much less sure about
    Rosita,}\\ {\bf and I suspect it is ok that Tim is not. Your call
    though. Also needed are updated affiliations for Francesco and
    Robert.}}

@@

% Abstract of the paper
\begin{abstract}
  Although they are intrinsically rare and not representative of the
  general population of radio galaxies, the most powerful radio
  galaxies are well studied, with a strong consensus regarding their
  formation processes. Despite being much more numerous, low-power
  radio galaxies are comparatively poorly studied, with many
  conflicting viewpoints regarding their origin and the importance of
  feedback within them. Here we present a complete sample of $11$
  nearby low-power radio galaxies, observed with a suite of
  instruments. We focus on the spatially-resolved stellar kinematics
  and populations, as well as the ionised gas distribution, kinematics
  and ionisation mechanisms derived from optical integral-field
  spectroscopic (IFS) observations.
  % , while other observation such as \ce{^{12}CO(2-1)} from Atacama
  % Large Millimetre/sub-millimetre Array (ALMA) and radio continuum
  % from the Very Large Array (VLA) will be presented in future papers.
  We compare the characteristics of our targets to those of ordinary
  (i.e.\ optically-selected) early-type galaxies from other IFS
  surveys.
  % using the Atlas$^\text{3D}$ and MASSIVE IFS surveys.
  Properties probed include the fast/slow rotator classification
  fraction, kinematic sub-structures, absorption line strength radial
  gradients, general stellar population properties, ionised gas
  distribution and kinematics, and the dominant sources of ionising
  photons. We tentatively find that our radio-selected early-types
  contain a higher mass of ionised gas than expected. However, after
  accounting for the differences in their mass distributions, we find
  no further difference between the different samples in terms of the
  characteristics that can be studied with optical IFSs. We suggest
  that this is consistent with the notion that low-power radio
  galaxies are ordinary early-type galaxies experiencing a brief
  active phase that maintains their low star-formation rates (also
  known as maintenance- or jet-mode active galactic nuclei).
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}

It is becoming increasingly clear that active galactic nuclei (AGN)
can be separated into two different types, radiative- and jet-mode AGN
\citep[e.g.][]{Antonucci2012, Heckman2014}, raising speculation about
differing mechanisms and possibly fuel sources for
each. Radiative-mode AGN are thought to be the powerful engines that
initially and rapidly shut off star formation in massive early-type
galaxies (ETGs; e.g.\ \citealt{Thomas2005, Thomas2010}), by either
removing the cold gas from the galaxy entirely, or by heating it such
that it joins the hot (X-ray) gas often present in massive galaxies
\citep[e.g.][]{OSullivan2001}.
% , or rendering it unusable for star formation (such as by
% stabilising it against collapse; e.g.\ \citealt{Martig2009}).
After a single radiative-mode episode, the AGN is thought to subside
to become jet-dominated, the jet repeatedly reheating the surrounding
interstellar medium (ISM) to counter any cooling that might allow the
cold gas reservoir to replenish (and subsequently form stars).

There is a strong consensus in the literature that the most powerful
radio galaxies are the result of gas-rich major mergers, that cause
large amounts of material to fall on to the black hole at the centre
of the resulting galaxy, thus producing an immensely powerful AGN that
includes a powerful radio jet \citep[e.g.][]{Malin1983, Quillen1992,
  Lim2000}. However, such galaxies are extremely rare and are not
representative of the general radio galaxy population. Most radio
galaxies are low powered, and their basic properties and origins are
poorly constrained. One of the most popular suggestions is that they
are bona-fide detections of jet-mode AGN \citep[e.g.][]{Heckman2014}.
% , a class of AGN thought to maintain the dearth of star formation in
% early-type galaxies (ETGs), after most of the fuel for star
% formation has been removed by an earlier event.
These low-power radio galaxies are known to be dusty, the dust lanes
having a tendency to be perpendicular to the jets
\citep[e.g.][]{DeRuiter2002, VerdoesKleijn2005}. They have little
ionised gas \citep[e.g.][]{Sarzi2005}, but often have significant
quantities of hot X-ray gas \citep[e.g.][]{Canizares1987}. The
molecular ISM of our sample galaxies has already been examined in
\citet{ruffa2019, ruffa2020}. If the low star-formation rates
typically seen in massive ETGs are preserved by jet-mode AGN, then all
these ETGs must regularly go through similarly active phases, and
therefore low-power radio galaxies should be much the same as their
radio-quiet (quiescent phase) ETG counterparts.

To test this hypothesis, a complete sample of $11$ nearby low-power
radio galaxies was constructed by \citet{Prandoni2010} and observed
with a suite of instruments. Here, we present the spatially-resolved
stellar kinematics and populations, as well as the ionised gas
distribution, kinematics and ionisation mechanisms derived from
integral-field spectroscopic (IFS) observations with the Visible
Multi-object Spectrograph (VIMOS) and archival data from the
Multi-unit Spectroscopic Explorer (MUSE), both on the Very Large
Telescope (VLT). We then compare the characteristics (accessible with
$V$-band IFS data) of our low-power radio galaxies to those of
ordinary (i.e.\ optically-selected) ETGs from the Atlas$^\text{3D}$
\citep{Cappellari2011} and MASSIVE \citep{Ma2014} IFS surveys. We
hereafter refer to the combined Atlas$^\text{3D}$ and MASSIVE sample
as the A+M sample.

The structure of this paper is as follows. We describe our complete
sample of low-power early-type radio galaxies in
Section~\ref{sec:samp}, and present our observations and data
reduction strategies in Section~\ref{sec:obs}. Our data analysis
methods and routines are detailed in Section~\ref{sec:analysis}, while
the various maps of our sample galaxies thereby produced are shown in
Appendix~\ref{sec:maps}. Our findings regarding the stars (kinematics,
absorption line strengths and populations) and ionised gas
(distribution, kinematics and ionisation mechanisms) are presented in
Sections~\ref{sec:StarKine}-\ref{sec:gas}. Our results are discussed
in Section~\ref{sec:discuss}, where we also conclude briefly.

\section{Sample Description}
\label{sec:samp}

The Parkes $2.7$~GHz survey \citep{Ekers1989} was used as a parent
sample from which to select our target galaxies, as it contains radio
sources (down to a flux density limit of $250$~mJy at $2.7$~GHz) with
an optical counterpart (with an integrated $V$-band apparent magnitude
$m_V\le17.0$). Host galaxies with a redshift greater than $0.03$ were
discarded. These criteria lead to a sample of $11$ galaxies, hereafter
referred to as our Southern Sample, whose key characteristics are
summarised in Table~\ref{tab:sample}. All galaxies have a FR~I radio
morphology according to the \citet{Fanaroff1974} scheme. It is worth
noting that both the radio and optical limits are based on apparent
fluxes, so this is not a volume-limited sample.

\begin{table*}[t]
  \begin{center}
    % \begin{threeparttable}
    \caption{Key characteristics of our Southern Sample galaxies.}
    \label{tab:sample}
    \begin{tabular*}{0.7\textwidth}{@{\extracolsep{\fill}}l c c c c l}
      \hline
      \hline
      Host galaxy & Radio source & Redshift & $S_\text{1.4\,GHz}$ & $m_K$ & Dust morphology\\
                  & (PKS) & & (mJy) & (mag) &\\
      \hline 
      ESO~443-G024 & 1258-321 & $0.01703$ & $1387\pm45$ & $8.51$ & no dust$^\text{a}$\\ 
      IC~1459 & 2254-367 & $0.00565$ & $1280\pm45$ & $6.81$ & dust lane$^\text{b}$\\
      IC~1531 & 0007-325 & $0.02553$ & $\leavevmode\phantom{0}388\pm12$ & $9.55$ & --\\
      IC~4296 & 1333-\leavevmode\phantom{0}33 & $0.01248$	& $1342\pm41$ & $7.50$ & edge-on disc$^\text{b}$\\
      NGC~612 & 0131-\leavevmode\phantom{0}36 & $0.02954$ & $\leavevmode\phantom{0}585\pm18$ & $9.58$ & dust lane$^\text{c}$\\
      NGC~1316 & 0320-\leavevmode\phantom{0}37 & $0.00591$ & $\leavevmode\phantom{0}255\pm10$ & $5.59$ & dust patches$^\text{b}$\\
      NGC~1399 & 0336-\leavevmode\phantom{0}35 & $0.00472$ & $\leavevmode\phantom{0}209\pm\leavevmode\phantom{0}7$ & $6.31$ & no dust$^\text{b}$\\
      NGC~3100 & 0958-314	& $0.00879$ & $\leavevmode\phantom{0}530\pm16$ & $8.08$ & dust lane$^\text{d}$\\
      NGC~3557 & 1107-372 & $0.01016$ & $\leavevmode\phantom{0}484\pm16$ & $7.20$ & face-on disc$^\text{b}$\\
      NGC~7075 & 2128-388 & $0.01819$ & $\leavevmode\phantom{0}837\pm28$ & $9.56$ & --\\
      -- & 0718-\leavevmode\phantom{0}34 & $0.02897$ & $1119\pm41$ & $9.97$ & dust patches$^\text{e}$\\
      \hline
      \hline
    \end{tabular*}
    % \begin{tablenotes}
    \parbox[t]{0.7\textwidth}{\footnotesize\textit{Notes:} Col.~1:
      galaxy name. Col.~2: radio source name from the Parkes Southern
      Radio Source Catalogue (PKS). We hereafter refer to the last
      galaxy by its PKS name only. Col.~3: redshift. Col.~4: National
      Radio Astronomy Observatory (NRAO) VLA Sky Survey (NVSS)
      $1.4$~GHz flux density \citep{Condon1998}. Col.~5: Two Micron
      All Sky Survey (2MASS) $K$-band apparent magnitude (with
      uncertainties of $\pm0.025$~mag;
      \citealt{Skrutskie2006}). Col.~6: dust morphology from (a)
      \citet{Govoni2000}, (b) \citet{Lauer2005}, (c)
      \citet{Bettoni2001}, (d) \citet{Sandage1979} and (e)
      \citet{Colbert2001}.}
    % \end{tablenotes}
    % \end{threeparttable}
  \end{center}
\end{table*}

Our Southern Sample was initially observed with the Atacama Pathfinder
Experiment \citep[APEX; ][]{Gusten2006}, with a detection claimed in
the \ce{^{12}CO(2-1)} transition for every galaxy
\citep{Prandoni2012}. Some galaxies had an extremely large line width
(up to a full width at half-maximum of $904$~km~s$^{-1}$ for IC~4296),
but most showed a flat or double-horn spectrum suggestive of ordered
rotation. Follow-up Atacama Large Millimeter/sub-millimeter Array
(ALMA) observations were recently presented in \citet{ruffa2019,
  ruffa2020}.

\section{Observing Strategy and Data Reduction}
\label{sec:obs}

\subsection{New VIMOS Data}
\label{subsec:VIMOS}

We have observed our Southern Sample targets with a suite of
instruments spanning radio to optical wavelengths. Archival X-ray
images from the \textit{Chandra space telescope} are also available
for some of the sample galaxies. This paper focuses on optical IFS
observations carried out with VIMOS, mounted on the VLT
\citep{LeFevre2003}. All observations were obtained with a spatial
sampling of $0\farcs67$, using the high-resolution blue grism yielding
a wavelength range of $3700$--$5520$~\AA\ with a spectral resolution
of $1440$ and a spectral sampling of $0.71$~\AA\ per pixel. Every
object was observed with a total integration time of $\approx100$~min
on-source, equally spread over three observing blocks (OBs). Each OB
also contained all necessary observations for standard IFS data
reduction.

VIMOS has several well-known though poorly-understood technical
problems. These include several low transmission (bad) fibres, strong
flexure, contamination from adjacent spectra on the charged coupled
device (CCD; cross-talk) and large differences of sensitivity across
its four quadrants. To address these issues, we adopted a
data-reduction pipeline built using \textsc{Py3D}, a suite of
programmes based on the \textsc{python} versions of the codes
developed for the Calar Alto Legacy Integral-field Spectroscopy Area
survey \citep[CALIFA;][]{Sanchez2012, Husemann2013}, but later updated
for VIMOS by \citet{Husemann2014}. \textsc{Py3D} applies all standard
IFS data-reductions steps: bias subtraction, spectrum identification,
flatfielding and wavelength calibration. The data were flux calibrated
using publicly available observations of the spectro-photometric
standard star Feige~110 provided by the European Southern Observatory
(ESO). All steps include procedures designed to account for the bad
fibres, flexure and cross-talk within VIMOS.

Following this, we noted that the cubes were still not properly
calibrated. Three main issues remained: the four quadrants had
inconsistent intensities (clearly showing uncorrected throughput
differences), diagonal intensity stripes were present in the
reconstructed images at all wavelengths, and spectral features were
observed that are visually similar to fringe patterns caused by
interference within the CCD (between incident light and light
reflected at the interfaces of the CCD layers;
\citealt{Jullo2008}). \citet{Lagerholm2012} presented ad-hoc methods
to correct for these issues, that we also adopt here. These additional
steps result in datacubes that are not perfectly flux-calibrated, but
all the corrections are multiplicative and will therefore not affect
equivalent width or line ratio measurements. From comparisons to MUSE
data of the same objects (see Section~\ref{subsec:MUSE}), we can
assess the flux calibration of the resulting datacubes.

Even these steps were unfortunately not totally effective, but we were
unable to further improve the data reduction. Some artefacts remain in
the data, but in most cases the data are of sufficient quality for our
purposes. For the worst affected datacubes, we fortunately have
alternative archival MUSE observations (see
Section~\ref{subsec:MUSE}).

The variance spectra are propagated throughout the data reduction
pipeline (including the aforementioned ad-hoc corrections) and are
square-rooted at this point, to be used as noise inputs in the
following analyses (see~Section \ref{sec:analysis}).

\subsection{MUSE Archival Data}
\label{subsec:MUSE}

Four of our Southern Sample galaxies have archival MUSE data analogous
to our VIMOS data: IC~1459, IC~4296, NGC~1316 and NGC~1399. We include
these data in our analysis as they have better spatial and spectral
resolutions and samplings, larger fields-of-view, larger wavelength
ranges, and better signal-to-noise ratios ($S/N$) than the VIMOS
data. There is also no significant issue with the data reduction.

We used the pre-reduced (Phase~3) datacubes from ESO. These were
generally of sufficient quality for our purpose, except for IC~1459
and IC~4296, for which the sky appears to have been
over-subtracted. To remedy this over-subtraction, we developed our own
pseudo-sky subtraction routine. For each galaxy, a median spectrum was
taken from four $20\times20$ spaxel$^2$ regions, one in each spatial
corner of the ESO reduced cube. After checking that no stellar
continuum could be fit to this median sky spectrum (i.e.\ that very
little galaxy light contaminates the pseudo-sky regions), this
spectrum was subtracted from each spaxel in the cube.

Finally, we trimmed all the MUSE cubes to the central
$30\arcsec\times30\arcsec$ ($150\times150$ spaxels$^2$) only, to (a)
exclude the regions used for pseudo-sky subtraction and (b) reduce the
computing resources required for spatial binning (see
Section~\ref{sec:analysis}). In any case, the bins would likely be so
large in the outer parts as to be effectively useless.

\section{Data Analysis}
\label{sec:analysis}

Both the VIMOS and MUSE datasets were spatially-binned using Voronoi
binning\footnote{http://www-astro.physics.ox.ac.uk/~mxc/software/}
\citep{Cappellari2003} and from this point on were treated almost
identically, the only difference (other than the different spectral
ranges and resolutions) being that the MUSE datacubes could and were
spatially-binned to higher $S/N$. For our purposes, we define the
`signal' and `noise' of each spaxel as the median value of its
spectrum and noise spectrum, respectively. For each bin, we require
$S/N\ge30$ for all VIMOS datacubes and $S/N\ge50$ for the IC~1459 and
IC~4296 MUSE datacubes. The NGC~1316 and NGC~1399 MUSE datacubes were
binned to $S/N\ge50$ for the stellar kinematics analysis and
$S/N\ge100$ for the the emission line kinematics and stellar
population analyses. These $S/N$ targets were chosen to be as low as
possible while still yielding useful maps (assessed by eye).
